---
title: "MFBVAR Alpha Verison"
author: "Caleb"
date: "2022-11-26"
output: html_document
---

```{r}
## load libraries
library(purrr)
library(tidyverse)
library(dbplyr)
library(mfbvar)
library(parallel)
```

## Load in Data 
```{r}
# load in data set from cleaning 
df <- read.csv("SCAN/mfvar.alpha.verison.csv")
``` 

## Transform Data 

You can also embed plots, for example:

```{r}
df.q <- read.csv("mfvar/quarterly.input.correct.csv")
df.m <- read.csv("mfvar/monthly.input.mfvar.pa.csv")

#join quarterly and monthly data 

df.mfvar <- full_join(df.m, df.q)

# select rows that you need, montly variables before quarterly ones 

df.alpha <- df.mfvar %>% 
  select("DATE", "ur", "manemp", "rgdp", "fpi")
# create log diff function

log_diff <- function(x, lag = 1) {c(rep(NA, lag), 1200/lag * diff(log(x), lag = lag))}

# turn variables into log diffs, keep unemployment rate as is, already rate 
 
df.alpha1 <- df.alpha %>% 
  mutate(ur = ur, 
         manemp = log_diff(manemp), 
         rgdp = log_diff(rgdp, lag = 3), 
         fpi = log_diff(fpi, lag = 3))

# need to start with a full row and end with a full row 

df.beta <- df.beta[-nrow(df.beta),]
# removes first n number of rows in c(n), 5 is specfic to monthly and quarterly data 
df.beta <- df.beta %>%  filter(!row_number() %in% c(1, 2, 3, 4, 5))
```
## Set Prior 
```{r} 

# set the prior, freq = c(put the frequency of every variable, going to be annoying for big data sets, not sure what n_burnin and n_reps does)
prior <- set_prior(Y = df.alpha2, freq = c("m", "m", "q", "q"), n_lags = 4, n_burnin = 1000, n_reps = 1000)
# plot initial prior 
plot(prior)

# print the prior, tells us what kinds of prior we can use 
prior 

```
## Calculate Steady-State Prior 
```{r} 
# get steady state means for the prior, I think this is just for a steady state prior, but not sure how we would do this for a minnesota style prior, not sure how to get the 95% confidence intervals in r, need sample mean and then standard erorr but defintely a function that can do it for me
n <- colSums(!is.na(df.alpha2))
n 
# calculate confidence interval for ur 
xbar.ur <- mean(df.alpha2$ur, na.rm = t) 
n.ur <- n[2]
sd.ur <- sd(df.alpha2$ur, na.rm = T) 
margin.ur <- qt(0.975,df=n.ur-1)*sd.ur/sqrt(n.ur)

lb.ur <- xbar.ur - margin.ur 

ub.ur <- xbar.ur + margin.ur 

ci.ur <- c(lb.ur, ub.ur)

ci.ur 

# this is really wide, I think the log differences got fucked up but we're just going to run with it 


#calculate confidence intveral for manemp 
xbar.manemp <- mean(df.alpha2$manemp, na.rm = T) 
n.manemp <- n[3]
sd.manemp <- sd(df.alpha2$manemp, na.rm = T) 
margin.manemp <- qt(0.975,df=n.manemp-1)*sd.manemp/sqrt(n.manemp)

lb.manemp <- xbar.manemp - margin.manemp

ub.manemp <- xbar.manemp + margin.manemp 

ci.manemp <- c(lb.manemp, ub.manemp)

ci.manemp


# CI for rgdp 
xbar.rgdp <- mean(df.alpha2$rgdp, na.rm = T) 
n.rgdp <- n[4]
sd.rgdp <- sd(df.alpha2$rgdp, na.rm = T) 
margin.rgdp <- qt(0.975,df=n.rgdp-1)*sd.rgdp/sqrt(n.rgdp)


lb.rgdp <- xbar.rgdp - margin.rgdp

ub.rgdp <- xbar.rgdp + margin.rgdp

ci.rgdp <- c(lb.rgdp, ub.rgdp)

ci.rgdp

# CI for fpi 
xbar.fpi <- mean(df.alpha2$fpi, na.rm = T) 
n.fpi <- n[5]
sd.fpi <- sd(df.alpha2$fpi, na.rm = T) 
margin.fpi <- qt(0.975,df=n.fpi-1)*sd.fpi/sqrt(n.fpi)

lb.fpi <- xbar.fpi - margin.fpi

ub.fpi <- xbar.fpi + margin.fpi

ci.fpi <- c(lb.fpi, ub.fpi)

ci.fpi

```
## Update Prior w/ Steady State 
```{r}

# update prior to include steady state data 
prior_intervals <- matrix(data = c(ci.ur,
                             ci.manemp, 
                             ci.rgdp, 
                             ci.fpi), ncol = 2,byrow = TRUE)

# print matrix to double check 

prior_intervals
# not sure what this does, assume it just converts it to a useable form  
moments <- interval_to_moments(prior_intervals)
prior <- update_prior(prior, d = "intercept",
                      prior_psi_mean = moments$prior_psi_mean,
                      prior_psi_Omega = moments$prior_psi_Omega)

#plot prior results 

plot(prior)



# update prior to specficy how long you want to forecast out 
prior <- update_prior(prior, n_fcst = 24)

#grid search for hyperparameters

#search for lambda 1 given lambda3 = 1
 par_fun <- function(lambda1, prior) {
  set.seed(2019)
  mod_par <- estimate_mfbvar(prior, prior = "ss", variance = "iw",
  lambda1 = lambda1, lambda3 = 1)
  mdd(mod_par)
 }

#run grid search
cl <- makeCluster(4)
clusterEvalQ(cl, library("mfbvar"))
lambda1_seq <- seq(0.05, 1, by = 0.05)
lambda1_result <- parSapply(cl, lambda1_seq,
  par_fun, prior = prior)
stopCluster(cl)

#update prior with new lambda1
prior <- update_prior(prior, lambda1 = lambda1_result)

#search for lambda 2 given lambda3 = 1
 par_fun <- function(lambda2, prior) {
  set.seed(2019)
  mod_par <- estimate_mfbvar(prior, prior = "ss", variance = "iw",
  lambda2 = lambda2, lambda3 = 1)
  mdd(mod_par)
 }

#run grid search
cl <- makeCluster(4)
clusterEvalQ(cl, library("mfbvar"))
lambda2_seq <- seq(0.05, 1, by = 0.05)
lambda2_result <- parSapply(cl, lambda1_seq,
  par_fun, prior = prior)
stopCluster(cl)

#update prior with new lambda1
prior <- update_prior(prior, lambda2 = lambda2_result)

# pull summary of prior 
summary(prior)

```

## Create prior with block exogenous variables 
```{r} 
prior.exo <- set_prior(Y = df.alpha2, freq = c("m", "m", "q", "q"), n_lags = 4, n_burnin = 1000, n_reps = 1000, block_exo = c("fpi", "manemp"), n_fcst = 24)
prior.exo
```
## Run model w/ Minnesota Prior and exogenous variables 
```{r} 
mod.minn.exo <- estimate_mfbvar(prior.exo, prior = "minn", variance = "diffuse")

predict(mod.minn.exo, pred_bands = .95)

plot(mod.minn.exo)
```
## Run model 

```{r} 

# run model with steady state prior
modv1.0 <- estimate_mfbvar(prior, prior = "ss", variance = "iw")

modv1.0 

```
## Get Results 
```{r}
# get results 

predict(modv1.0, pred_bands = NULL)

predict(modv1.0, aggregate_fcst = TRUE, pred_bands = .8)
```

## Min Prior Model 
```{r} 
# run model with minn prior 
mod.minn.iw <- estimate_mfbvar(prior, prior = "minn", variance = "iw")

mod.minn.iw


```

## Get Results 
```{r} 
# predict model with 95% confidence intervals 
df.mod.minn.iw95 <- predict(mod.minn.iw, aggregate_fcst = TRUE, pred_bands = .95)

df.mod.minn.iw95

df.mod.minn.iw95 <- as.data.frame(df.mod.minn.iw95)

# predict model with 80% confidence intervals

df.mod.minn.iw <- predict(mod.minn.iw, aggregate_fcst = TRUE, pred_bands = .8)

df.mod.minn.iw

df.mod.minn.iw <- as.data.frame(df.mod.minn.iw)
```
## Try to plot minn model 
```{r} 
plot(mod.minn.iw, plot_start = "2005-08-01", nrow_facet = 3)

```
## Get MDD for model w/ steady state prior 
```{r} 
mdd(mod_ss_iw)
```

## Get MDD for model w/ Minnesota Prior 
```{r} 
#p_trunc giving the degree of truncation of the truncated normal distribution, needed when run model w/ a minnesota prior 
mdd(mod.minn.iw, p_trunc = 1)
```